---
title: "Data Preparation"
description: "Load and format modeling data for analysis"
format:
  html: 
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options: 
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```

## Load Libraries  
Load needed R-packages to execute code.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(usefun)
library(ggplot2)
library(here)

select <- dplyr::select
```

## Custom Functions  
Custom functions to process ISP outputs and generate figure plots.  These are available from the linked repository at the top of the script.  
```{r message=FALSE, warning=FALSE}
source(here("./R/utilities.R"))
source_dir("./R")
```

## Import Data
The `import_infection_data` function reads a `.txt` file containing infection events (`Outputs_Infection.txt`) for a specified region assigned to `regional_scenario`.

### Assign File Path
```{r}
generic_path <- "./script-inputs/isp-outputs"

## Assign regional scenario
regional_scenario <- "eastern"

regional_path <- sprintf("%s/outputs_%s", generic_path, regional_scenario)
```


### Import infection data
```{r}
infection <- import_infection_data(regional_path)

dim(infection)
```