---
title: "Data Preparation"
description: "Load and format modeling data for analysis"
format:
  html: 
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options: 
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```

## Load Libraries  
Load needed R-packages to execute code.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(usefun)
library(ggplot2)
library(here)
library(patchwork)
library(scales)
library(geosphere)
library(terra)
library(raster)
library(sf)
library(viridis)
library(pals)

select <- dplyr::select
```

## Custom Functions  
Custom functions to process ISP outputs and generate figure plots.  These are available from the linked repository at the top of the script.  
```{r message=FALSE, warning=FALSE}
source(here("./R/utilities.R"))
source_dir("./R")
```

## Assign File Path
```{r}
generic_path <- "./script-inputs/isp-outputs"

## Assign regional scenario
regional_scenario <- "eastern"

regional_path <- sprintf("%s/outputs_%s", generic_path, regional_scenario)
```

## Import Data

### Import Infection Data
The `import_infection_data` function reads a `.txt` file containing infection events (`Outputs_Infection.txt`) for a specified region.
```{r}
infection <- import_infection_data(regional_path)

dim(infection)
```

### Import Detection Data
The `import_detection_data` function reads a `.txt` file containing detection events (`Outputs_Detection.txt`) for a specified region.
```{r}
detection <- import_detection_data(regional_path)

dim(detection)
```

## Read Farm Reference File
The FMD Farm Reference File used for this project contains geographic coordinates for theoretical farm locations. Premises data were generated using the Farm Location and Agricultural Production Simulator (FLAPS, see Burdett et al. 2015) and do not reflect true farm locations. Farm demographics are based upon National Agricultural Statistical Service (NASS) Census of Agriculture data.
```{r}
reference <- read.csv2(here("script-inputs/reference-files/FMD_Reference_File.csv"), header=TRUE, sep=",")

colnames(reference) <- c(
                      "premises",
                      "class",
                      "cattle",
                      "weight",
                      "x",
                      "y",
                      "state",
                      "longitude",
                      "latitude",
                      "geo_region"
                      )

## Select columns needed for spread analyses
reference <- reference %>%
  mutate(longitude = as.numeric(longitude),
         latitude = as.numeric(latitude)) %>%
  select(-c(weight, cattle, x, y, geo_region))

dim(reference)
```