---
title: "Epidemic Duration"
description: "Summarize the epidemic duration of simulated outbreaks"
format:
  html: 
    df-print: kable
    code-fold: true
    code-summary: "Code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options: 
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
options(dplyr.summarise.inform = FALSE)

library(usefun)
library(ggplot2)
library(here)
library(patchwork)
library(scales)

## Custom Functions
source(here("./R/utilities.R"))
source_dir("./R")

## Assign file path 
generic_path <- "./script-inputs/isp-outputs"

## Assign regional scenario
regional_scenario <- "eastern"

regional_path <- sprintf("%s/outputs_%s", generic_path, regional_scenario)

## Import detection data
detection <- import_detection_data(regional_path)
```

## Detection Summary
The `summarize_detections` function generates a data frame summarizing detection events by iteration. The `generate_detect_statistics` generates a data frame of summary statistics grouped by preclinical infectious duration and response type.
```{r}
detect_summary <- summarize_detections(detection)

head(detect_summary, 3)

detect_config_summary <- generate_detect_statistics(detect_summary)

head(detect_config_summary, 3)
```

## Epidemic Duration
The *epidemic duration* is calculated as the number of days between the first detection and last detection of disease in a single outbreak.
```{r}
duration_summary <- detect_config_summary %>%         
  filter(summary == "duration")

dim(duration_summary)

## Print summary of Epidemic Duration (days)
duration_summary
```

## Significance Test
Significance testing performed on optimal and suboptimal detection scenarios
```{r}
## Filter out low virulence scenarios
no_low_virulence <- detection %>%
  filter(response_type != "low-virulence")

no_low_virulence_summary <- summarize_detections(no_low_virulence)

range(no_low_virulence_summary$duration)
plot(density(no_low_virulence_summary$duration))
```

### Linear Model
```{r}
model_duration <- lm(duration ~ preclinical * response_type, 
                       data = no_low_virulence_summary)

summary(model_duration)
```

### Analysis of Variance (ANOVA)
```{r}
anova(model_duration)
```

## Plot Duration
Prepare to plot the calculated median epidemic duration (days)
```{r}
duration_summary$response_type <- ordered(factor(duration_summary$response_type), c("optimal", "suboptimal", "low-virulence"))

response_colors <- c("#74add1", "orange2", "red4") 
```

### Generate Plot Using Patchwork Package
The `plot_epidemic_duration` function plots epidemic duration by scenario, with epidemic duration (days) on the x-axis and the duration of incubation phase transmission on the y-axis.
```{r, message = FALSE, warning = FALSE}
duration_plot <- plot_epidemic_duration(duration_summary, regional_scenario)
```

```{r fig.width=12, fig.height=10, message=FALSE, warning=FALSE}
duration_plot
```