---
title: "Overview"
description: "Analyses of ISP outputs from a FMDV preclinical transmission model in U.S. cattle"
format:
  html: 
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options: 
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```

::: {.callout-important icon=true}
#### Under Active Development  

This site includes exploratory analyses that have not been verified or validated and are subject to change. 
  
:::
  
  
### Navigation  
The menu at right can be used to navigate through major sections on a webpage.  Links across the top of the page will direct to specific analyses.

## Load Libraries  
Load needed R-packages to execute code.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(usefun)
library(ggplot2)
library(here)

select <- dplyr::select
```

## Custom Functions  
Custom functions to process ISP outputs and generate figure plots.  These are available from the linked repository at the top of the script.  
```{r message=FALSE, warning=FALSE}
source(here("./R/utilities.R"))
source_dir("./R")
```

## Assign File Path
```{r}
generic_path <- "./script-inputs/isp-outputs"
```
